A Horrocks-type theorem for even orthogonal $K_2$

### 2020-03-26 dl-paper.tex
Fixing issues found by myself

Throughout this section we adopt the following notation:
\begin{itemize}
 \item $A$ is an arbitarary commutative ring and $M$ is an ideal of $A$. 
 In Section~\ref{sec:presentation} there are no restrictions on $(A, M)$, but in later sections the pair $(A, M)$ is assumed to be local.
 \item $R$ denotes the Laurent polynomial ring $A[X, X^{-1}]$. It will be convenient for us to set $t = X^{-1}$, so also $R = A[t, t^{-1}]$.
 \item $B$ denotes the subring $A[t] + M[t^{-1}]$ of $R$ consisting of Laurent polynomials $f(t,t^{-1})$ whose coefficients of terms of negative degree belong to $M$;
 \item $I$ denotes the ideal $M[t, t^{-1}]$ of $R$. $I$ can also be considered as an ideal of $B$.
\end{itemize}
\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
Let $M\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$.
We denote by $B = B(A, M)$ the subring $A[t] + M[t\inv] \subseteq A[t, t\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. Obviously, $B = A[t]$ in the case $M=0$ and $B = A[t, t\inv]$ in the case $M=A$.

Whenever the coefficient $\xi$ of a Steinberg generator $g = x_\alpha(\xi)$ of $\St(\Phi, B)$ is a homogeneous element of $B$, i.\,e. $\xi \in B_d$ for some $d \in \mathbb{Z}$,
 we call the corresponding generator $g$ {\it homogeneous of degree $d$}.
It is not hard to show that $\St(\Phi, B)$ can be presented by the set of all homogeneous Steinberg generators modulo the following set of Steinberg relations
(below $a, a' \in B_d$, $b\in R_e$ and $d,e \in \mathbb{Z}$): 
\begin{align}
&\,\,x_{\alpha}(a)\cdot x_{\alpha}(a') =  x_{\alpha}(a+a'),                        & \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),   & \alpha + \beta \in \Phi,\ \tag{R$2_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha \perp \beta.\ \tag{R$3^\bot_{d,e}$}
\end{align}
By the degree of a Steinberg relation we mean the maximum of degrees of generators that appear in the relation.
For example, the degree of every relation of type $\text{R2}_{d,e}$ is $\max(d,e,d+e)$ 
 while the degree of a relation of type $\text{R3}^\bot_{d,e}$ or $\text{R3}^\angle_{d,e}$ is $\max(d,e)$.

For $n\geq 1$ we define the ``truncated'' Steinberg group $\St^{\leq n}(\Phi, B)$ by means of the set $\mathcal{X}_{\leq n}^\Phi$ of homogeneous Steinberg generators of degree $\leq n$ and the set of Steinberg relations $\mathcal{R}_{\leq n}^\Phi$ of degree $\leq n$. We denote by $F(\mathcal{X}^\Phi_{\leq n})$ the free group on $\mathcal{X}^\Phi_{\leq n}$.

>>>>

Throughout this section we use the following notations and conventions:
\begin{itemize}
 \item Let $A$ be an arbitarary commutative ring and $M$ be an ideal of $A$. Starting from subsection~\ref{sec:Pa0-basic} we also assume that $A$ is local and $M$ is the maximal ideal of $A$.
 \item We denote by $R$ the Laurent polynomial ring $A[X, X^{-1}]$. We set $t = X^{-1}$, so that $R = A[X, X^{-1}] = A[t, t^{-1}]$.
 \item $B = B_{A, M}$ denotes the subring $A[t] + M[t^{-1}] = A[X^{-1}] + M[X]$ of $R$. Obviously, $B = A[t]$ in the case $M=0$ and $B = R$ in the case $M=A$.
 \end{itemize}
\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
We consider $R = A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$. This grading induces the grading on the subring $B \subseteq R$. For an integer $d$ we denote by $R_d$ (resp. $B_d$) the degree $d$ part of the ring $R$ (resp. $B$). Obviously, $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. With this notation, $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ as an $A$-module.

Whenever the coefficient $\xi$ of a Steinberg generator $g = x_\alpha(\xi)$ of $\St(\Phi, B)$ is a homogeneous element of $B$, i.\,e. $\xi \in B_d$ for some $d \in \mathbb{Z}$,
 we call the corresponding generator $g$ {\it homogeneous of degree $d$}.
We denote by $\St^h(\Phi, B)$ the Steinberg group in the homogeneous presentation, i.\,e. the group given by the set of all homogeneous Steinberg generators modulo the following set of relations which we call homogeneous Steinberg relations
(below $a, a' \in B_d$, $b\in R_e$ and $d,e \in \mathbb{Z}$): 
\begin{align}
&\,\,x_{\alpha}(a)\cdot x_{\alpha}(a') =  x_{\alpha}(a+a'),                        & \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),   & \alpha + \beta \in \Phi,\ \tag{R$2_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha \perp \beta.\ \tag{R$3^\bot_{d,e}$}
\end{align}
\begin{rem}
Notice that in the above presentation we omitted the relations $[x_\alpha(a), x_\alpha(b)] = 1$ but it easy to see that they, in fact, follow from~\eqref{eq:H1iii},~$\text{R2}_{d,e}$ and $R3_{d,e}^\angle$: \[ ???. \]
\end{rem}

It is not hard to show that the map $\St^h(\Phi, B) \to \St(\Phi, B)$ induced by the obvious embedding of generators is an isomorphism.

By the degree of a Steinberg relation we mean the maximum of degrees of generators that appear in the relation.
For example, the degree of every relation of type $\text{R2}_{d,e}$ is $\max(d,e,d+e)$ 
 while the degree of a relation of type $\text{R3}^\bot_{d,e}$ or $\text{R3}^\angle_{d,e}$ is $\max(d,e)$.

For $n\geq 1$ we define the ``truncated'' Steinberg group $\St^{\leq n}(\Phi, B)$ by means of the set $\mathcal{X}_{\leq n}^\Phi$ of homogeneous Steinberg generators of degree $\leq n$ and the set of homogeneous Steinberg relations $\mathcal{R}_{\leq n}^\Phi$ of degree $\leq n$.  We denote by $F(\mathcal{X}^\Phi_{\leq n})$ the free group on $\mathcal{X}^\Phi_{\leq n}$.

Notice that $\lim\limits_{n \to +\infty} \St^{\leq n}(\Phi, B) = \St^h(\Phi, B) \cong \St(\Phi, B)$.
